---
title: 'Health Data Study with Visualization'
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

Reading training and testing dataset available in csv format  

```{r}
training_data = read.csv(file = 'ALS_TrainingData_2223.csv')
testing_data = read.csv(file = 'ALS_TestingData_78.csv')
```
  
  
First 6 rows of the training dataset  

```{r}
head(training_data)
```
  
  
Size of the training dataset  
  
```{r}
print(paste("No of columns in training data: ", ncol(training_data)))
print(paste("No of rows in training data: ", nrow(training_data)))
```
  
  
Summary statistics of the training dataset
  
```{r}
summary(training_data)
```
  
  
First 6 rows of the testing data
  
```{r}
head(testing_data)
```
  
  
Size of the testing dataset  
  
```{r}
print(paste("No of columns in testing data: ", ncol(testing_data)))
print(paste("No of rows in testing data: ", nrow(testing_data)))
```
  
  
Summary statistics of the testing data  
  
```{r}
summary(testing_data)
```
  
  
Histograms are a visual representation of a dataset.  
It consists of x-axis, y-axis and various bars of different heights, where y-axis shows how frequently the values on x-axis occur in data.  
  
Histogram of the age data in training dataset as below shows more data for the age between the range 45 to 65  
  
```{r}
hist(training_data$Age_mean, col = 'aliceblue', main = 'Histogram of Age in training data', xlab = 'Age')
```
  
  
Histogram of the age data in testing dataset as below shows more data for the age between the range 45 to 65  
  
```{r}
hist(testing_data$Age_mean, col = 'darkturquoise', main = 'Histogram of Age in testing data', xlab = 'Age')
```
  
  
Majority of the data in training dataset has count of Albumin_median falling between 40 to 47  
  
```{r}
hist(training_data$Albumin_median, col = 'aliceblue', main = 'Histogram of Albumin in training data', xlab = 'Albumin_median')
```
  
   
Majority of the data in training dataset has count of ALSFRS_Total_median falling between 15 to 35  
  
```{r}
hist(training_data$ALSFRS_Total_median, col = 'aliceblue', main = 'Histogram of ALSFRS in training data', xlab = 'ALSFRS_Total_median')
```
  
  
Majority of the data in training dataset has count of Calcium_median falling between 2.2 to 2.5  
  
```{r}
hist(training_data$Calcium_median, col = 'aliceblue', main = 'Histogram of Calcium in training data', xlab = 'Calcium_median')
```
  
  
Density plots are used to view the distribution of a variable in a data  
  
Below graph shows the density data of 'White.Blood.Cell..WBC._median' in testing dataset.  
  
```{r}
d <- density(testing_data$White.Blood.Cell..WBC._median)
plot(d, main = "Density of 'White Blood Cell' variable in testing data")
polygon(d, col="darkturquoise")
```
  
  
Heatmap is used to visualize hierarchical clustering, where data values are transformed to color scale.  
  
First hierarchical clustering is done on both rows and columns of the data matrix, then data is re-ordered according to the hierarchical result, putting similar observations close to each other.  
  
heatmap() is a base function in R to draw a simple heatmap, where the 'scale' attribute indicates if the values should be centered and scaled in either row or column direction, or none.  
  
Below Heatmap of the training data  
  
```{r}
df <- scale(training_data)
heatmap(df, scale = "row")
```
  
  
Boxplots are used to measure how well the data is distributed in the dataset.  
It divides the data in three quartiles, the graph represents min, max, median, first quartile, third quartile in the dataset.  
  
```{r}
boxplot(training_data$Hemoglobin_median, xlab = "Hemoglobin_median", main = "Measurement of the Hemoglobin_median")
```
  
  
The below script will create a boxplot graph for the relation between pulse_median and bp_diastolic_median  
  
```{r}
boxplot(pulse_median ~ bp_diastolic_median, data = training_data, xlab = "bp_diastolic_median", ylab = "pulse_median", main = "Relationship between bp_diastolic and pulse")
```
