---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidytext)
```
```{r}
nrow(sentiments)
```
```{r}
sentiments
```
```{r}
get_sentiments("bing")
```

#### janeaustenr: An R Package for Jane Austen's Complete Novels
#### austen_books: Tidy data frame of Jane Austen's 6 completed, published novels

```{r}
library(janeaustenr)
```

```{r}
austen_books() 
```

#### Calculate Cumulative Sum of a Numeric Object in R Programming – cumsum() Function.
#### The cumulative sum can be defined as the sum of a set of numbers as the sum value grows with the sequence of numbers. 
```{r}
library(dplyr)
```
```{r}
library(stringr)
```

```{r}
tidy_data <- austen_books() %>% group_by(book) %>% 
  mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) %>% ungroup() %>% unnest_tokens(word, text)
```
```{r}
positive_senti <- get_sentiments("bing") %>% filter(sentiment == "positive")
positive_senti
```


```{r}
tidy_data %>% filter(book == "Emma") %>% semi_join(positive_senti) %>% count(word, sort = TRUE)
```
```{r}
library(tidyr)
bing <- get_sentiments("bing")
Emma_sentiment <- tidy_data %>%
 inner_join(bing) %>%
 count(book = "Emma" , index = linenumber %/% 80, sentiment) %>%
 spread(sentiment, n, fill = 0) %>%
 mutate(sentiment = positive - negative)
```
```{r}
library(ggplot2)
```
```{r}
ggplot(Emma_sentiment, aes(index, sentiment, fill = book)) +
 geom_bar(stat = "identity", show.legend = TRUE) +
 facet_wrap(~book, ncol = 2, scales = "free_x")
```
```{r}
counting_words <- tidy_data %>%
 inner_join(bing) %>%
 count(word, sentiment, sort = TRUE)
head(counting_words)
```
```{r}
counting_words %>%
 filter(n > 150) %>%
 mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
 mutate(word = reorder(word, n)) %>%
 ggplot(aes(word, n, fill = sentiment))+
 geom_col() +
 coord_flip() +
 labs(y = "Sentiment Score")
```
```{r}
#install.packages("wordcloud")
```


```{r}
library(reshape2)
library(wordcloud)
tidy_data %>%
 inner_join(bing) %>%
 count(word, sentiment, sort = TRUE) %>%
 acast(word ~ sentiment, value.var = "n", fill = 0) %>%
 comparison.cloud(colors = c("red", "dark green"),
          max.words = 100)
```

#### In this project, we went through our project of sentiment analysis in R.
#### We learnt about the concept of sentiment analysis and implemented it over the dataset of Jane Austen’s books. 
#### We were able to delineate it through various visualizations after we performed data wrangling on our data. 
#### We used a lexical analyzer – ‘bing’ in this instance of our project. Furthermore, we also represented the sentiment score through a plot and also made a visual report of wordcloud.
