---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

# Predicting number of Covid19 deaths using Time Series Analysis (ARIMA MODEL)


### Importing required libraries

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(zoo))
suppressMessages(library(aTSA))
suppressMessages(library(tseries))
suppressMessages(library(forecast))
suppressMessages(library(lubridate))
```

### Load the dataset

```{r}
covid_data <- read.csv("owid-covid-data.csv")
colnames(covid_data)
```

```{r}
str(covid_data)
```

### We change the date column into a date type

```{r}
covid_data$date <- as.Date(covid_data$date)
```

### We pick the US rows only.

```{r}
data_US <- covid_data[covid_data$location == "United States",c(3,4,7,8)]

summary(data_US)
```

### Checking for null or na rows

```{r}
sum(is.null(data_US$date))
sum(is.na(data_US$date))
sum(is.na(data_US$total_deaths))
sum(is.null(data_US$total_deaths))
sum(is.na(data_US$new_deaths))
sum(is.na(data_US$new_deaths))
```

### Plotting the number of deaths in the US

```{r}
p <- ggplot(data_US, aes(x=date, y=new_deaths)) +
  geom_line() + scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  xlab("Months") + ylab("New Deaths")+ggtitle("Number of deaths in the US")+geom_area(fill="lightblue", color="black")
p
```

### Plotting the mean to check if our data is stationary or not

```{r}
p + geom_smooth(method = lm, col = "red", se = FALSE)
```

### General trend of our data

```{r}
p1 <- ggplot(data_US, aes(x=date, y=total_deaths)) +
  geom_line() + scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  xlab("Months") + ylab("Aggregate Number of Deaths")+ggtitle("Trend Line")+geom_area(fill="red", color="black")
p1

```

### Here we can see the aggregated numbers of deaths

```{r}
data_US$total_deaths
```

### Change the data to zoo type providing the sequence as dates ranging from 2019/12/31 to 2020/08/01

```{r}
data_US_Analysis <- data_US[,c(2,4)]
data_demo <- zoo(data_US_Analysis$new_deaths, seq(from = as.Date("2019-12-31"), to = as.Date("2020-08-01"), by = 1))

summary(data_demo)

```

### Plotting the dataframe

```{r}
plot((data_demo))
```

### Converting the data to stationary by using diff function

```{r}
stationary_data <- diff(data_demo)
plot(stationary_data)
```

### To check if it's stationary we conduct a quantitative test. We use the Augmented Dickey-Fuller Test.
#### H_0 = The null hypothesis for this test is that there is a unit root.
#### H_A = The alternative hypothesis is that the time series is stationary (or trend-stationary).


### We select a significance level of 0.05 and since our p-value is 0.01 and smaller then 0.05, we come to the conclusion to reject the null hypothesis.

```{r}
adf.test(as.matrix(stationary_data)) 
```


## Modeling

### We use the Auto Correlation Graph when our data isn't stationary.
#### Here we see that our values are all exceeding the blue line. The goal is to have the values under the blue line and they should be inverted as well.

```{r}
acf(data_demo)
```

### To select the p and q values we select the number before the first inverted line

```{r}
acf(stationary_data)
```

```{r}
pacf(stationary_data)
```

### arima has a auto.arima function which gives us the ideal arima model based on our data.

```{r}
arima_funct <- auto.arima(stationary_data)
arima_funct
```

### lets use the auto.arima function to forecast 3 weeks

```{r}
forecast1 <- forecast(arima_funct, h=21)
additional_deaths <- round(sum(forecast1$upper[,2]))
# [1] 18589

total_number_of_deaths <- round(sum(data_US_Analysis$new_deaths)+additional_deaths)
# [1] 171903

additional_deaths
total_number_of_deaths
forecast1
```
```{r}
forecast1$upper[, 2]
```


```{r}
plot(forecast1)
```


### lets use the auto.arima function to forecast 3 months

```{r}
forecast2 <- forecast(object = arima_funct, h = 90)
additional_deaths2 <- round(sum(forecast2$upper[,2]),0)
# [1] 82653

total_number_of_deaths2 <- round(sum(data_US_Analysis$new_deaths)+additional_deaths2,0)
# [1] 235967

# additional_deaths2
# total_number_of_deaths2
forecast2
```

```{r}
plot(forecast2)
```


### Below we are just adding the dates and make changes to the x axis for the projected graphs

```{r}
delta <- (forecast1$lower[, 2] + forecast1$upper[, 2])/2

predicted_21 <- data.frame(date = seq(from = 18476, by = 1, length=21), new_deaths = delta)
original <- data.frame(date = as.numeric(data_US_Analysis$date), new_deaths = c(0, stationary_data))

z <- rbind(original,predicted_21)
z$date <- as.Date(z$date)

z
```

```{r}
forecast1_data <- as.data.frame(forecast1)
forecast1_data <- data.frame(date=seq(from=18476, by=1, length=21), forecast1_data, y = delta)
forecast1_data$date <- as.Date(forecast1_data$date)

graph <- ggplot(data=z,aes(x=date,y=new_deaths),show.legend = FALSE)+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  geom_line(data=original,aes(x=as.Date(date),y=new_deaths))+
  geom_ribbon(data=forecast1_data,aes(x = date,ymin =Lo.95, ymax =Hi.95), inherit.aes = FALSE,fill = "lightsteelblue2")+
  geom_ribbon(data=forecast1_data,aes(x = date,ymin =Lo.80, ymax =Hi.80), inherit.aes = FALSE,fill = "lightsteelblue3")+
  geom_line(data=forecast1_data,aes(x=date,y=y),size=1,color='purple')+
  ggtitle("Forecasts from ARIMA(3,0,2) with zero mean")+
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y")+
  labs(x="Date",y="New Deaths")

graph
```

#### Here, the forecasts are shown as a purple line, with the 80% prediction intervals as a dark shaded area and the 95% prediction intervals as a light shaded area.

```{r}
delta2 <- (forecast2$lower[,2]+forecast2$upper[,2])/2

predicted_22 <- data.frame(date=seq(from=18476,by=1,length=nrow(forecast2$lower)),new_deaths=delta2) # prediction from ARIMA
original <- data.frame(date=as.numeric(data_US_Analysis$date),new_deaths=c(0,stationary_data)) # Previous Observations 
z2 <- rbind(original,predicted_22) # combine the two data sets
z2$date <- as.Date(z2$date) # convert numeric column to date (creating the base layer of ggplot)
z2
```


```{r}
forecast2_data <- as.data.frame(forecast2) # convert forecast data into data.frame
forecast2_data <- data.frame(date=seq(from=18476,by=1,length=90),forecast2_data, y = delta2) #add a date and delta of Lo.95 & Hi.95
forecast2_data$date <- as.Date(forecast2_data$date) # covert to date 

graph1 <- ggplot(data=z2,aes(x=date,y=new_deaths),show.legend = FALSE)+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  geom_line(data=original,aes(x=as.Date(date),y=new_deaths))+
  geom_ribbon(data=forecast2_data,aes(x = date,ymin =Lo.95, ymax =Hi.95), inherit.aes = FALSE,fill = "lightsteelblue2")+
  geom_ribbon(data=forecast2_data,aes(x = date,ymin =Lo.80, ymax =Hi.80), inherit.aes = FALSE,fill = "lightsteelblue3")+
  geom_line(data=forecast2_data,aes(x=date,y=y),size=1,color='purple')+
  ggtitle("Forecasts from ARIMA(3,0,2) with zero mean")+
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y")+
  labs(x="Date",y="New Deaths")

graph1
```


```{r}
covid_new_data <- read.csv("owid-covid-data1.csv")

covid_new_data$date <- as.Date(covid_new_data$date)
```

```{r}
new_data_US <- covid_new_data[covid_new_data$location == "United States",c(3,4,8,9)]

new_data_US <- new_data_US %>%
  filter(date < '2020-11-02')

new_data_US[is.na(new_data_US)] <- 0
#head(new_data_US)
```


```{r}
p2 <- ggplot(new_data_US, aes(x=date, y=total_deaths)) +
  geom_line() + scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  xlab("Months") + ylab("Aggregate Number of Deaths")+ggtitle("Trend Line")+geom_area(fill="red", color="black")
p2

```

### Comparing with actual records

```{r}
temp_21 <- new_data_US %>% 
  filter(date > '2020-08-01' & date < '2020-08-22')

additional_21 <- round(sum(temp_21$new_deaths))
# [1] 18589

total_21 <- round(sum(data_US_Analysis$new_deaths) + additional_21)
# [1] 171903

additional_21
total_21

print(paste0("Predicted deaths in next 21 days: ", additional_deaths))
print(paste0("Predicted total deaths in next 21 days: ", total_number_of_deaths))

print(paste0("Difference with actual additional deaths: ", abs(additional_deaths - additional_21)))
print(paste0("Difference with actual total deaths: ", abs(total_number_of_deaths - total_21)))
```


```{r}
temp_90 <- new_data_US %>% 
  filter(date > '2020-08-01' & date < '2020-11-01')

additional_90 <- round(sum(temp_90$new_deaths))
# [1] 82653

total_90 <- round(sum(data_US_Analysis$new_deaths) + additional_90)
# [1] 235967

additional_90
total_90

# additional_deaths2
# total_number_of_deaths2

print(paste0("Predicted deaths in next 90 days: ", additional_deaths2))
print(paste0("Predicted total deaths in next 90 days: ", total_number_of_deaths2))

print(paste0("Difference with actual additional deaths: ", abs(additional_deaths2 - additional_90)))
print(paste0("Difference with actual total deaths: ", abs(total_number_of_deaths2 - total_90)))

```








