---
title: "Customer Segmentation"
output:
  word_document: default
  html_notebook: default
---
# Final Project
# Subject: Marketing Analytics
# Student Name: Tejal Parmar
# Student ID: 0802653
# Project Title: Customer Segmentation using R


### Importing required libraries

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library("repr"))
suppressMessages(library(stats))
suppressMessages(library("dplyr"))
suppressMessages(library("ggplot2"))
suppressMessages(library("scales"))
suppressMessages(library("lubridate"))
suppressMessages(library("ggcorrplot"))
suppressMessages(library("cohorts"))
suppressMessages(library("rfm"))
suppressMessages(library("gridExtra"))
```

## Data Importing and Visualization
### Loading dataset from csv file and summarizing the dataset

```{r}
df <- read.csv("SuperStoreOrders.csv")
head(df)
```

### Looking at structure of the dataset

```{r}
str(df)
```

## Data Cleaning
### Removing NA records

```{r}
df <- na.omit(df)
print(paste0("After removing NAs ", nrow(df), " records left"))
```

### Removing duplicate records using unique function

```{r}
df <- unique(df)
print(paste0("After removing duplicates ", nrow(df), " records left in the dataset."))
```

### Selecting project natives from the dataset for the project and have a look at the dataset

```{r}
ProjectNatives <- c("Continent", "Country_Region", "Region", "State", "City", "Customer_Segment", "Department", "Category", "Customer_Id", "Customer_Name", "Order_Id", "Order_Date", "Order_Priority", "Item", "Container", "Ship_Date", "Ship_Mode", "Discount", "Order_Quantity", "Profit", "Sales", "Shipping_Cost", "Unit_Price")

store_data <- df[ProjectNatives]

head(store_data)
```

### Removing item names containing only digits

```{r}
store_data <- store_data %>%
  filter(!(str_detect(store_data$Item, pattern = "^[:digit:]+$")))

print(paste0("Total number of records left in the dataset: ", nrow(store_data)))
```

### Extracting first word from the product name to populate Brand as a new column

```{r}
store_data <- store_data %>%
  mutate(Brand = str_extract(store_data$Item, "(\\w+)"))

unique(store_data$Brand)
```


```{r}
store_data <- store_data %>%
  mutate(Total_Price = Order_Quantity*Unit_Price)

head(store_data)
```

### Replacing brand names containing only digits with "Unknown"

```{r}
store_data$Brand <- str_replace(store_data$Brand, "^[:digit:]+$", "Unknown")

print(paste0("Total number of unique items in the dataset: ", length(unique(store_data$Item))))
print(paste0("Total number of unique brands in the dataset: ", length(unique(store_data$Brand))))
```

```{r}
# write.csv(store_data,"C:/Users/19054/Documents/Sem-3/303/Project/store_data_selected.csv", row.names = FALSE)
```


### Sales comparision in different Continents
```{r}
#Aggregating data by 'Continent' and Finding sum of 'Sales' 
Continent_Sales<- aggregate(Sales ~ Continent, data = store_data, sum)

#Changing column name of sales 
colnames(Continent_Sales)[2] <- "Total_Sales"

#Finding out Store with highest Sales 

Continent_Sales <-arrange(Continent_Sales, desc(Total_Sales)) #Arranged Continents based on Sales in descending order

# Converting Continent column into factor so that order won't change for graph 
Continent_Sales$Continent <- factor(Continent_Sales$Continent, levels = unique(Continent_Sales$Continent))

#Plotting Continent vs TotalSales

ggplot(data = Continent_Sales, aes(x = Continent, y = Total_Sales)) + 
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 13)) + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) + ggtitle('Continents vs Sales') +
  xlab("Continents") + ylab("Total Sales")

```

### Sales comparision in different Countries

```{r}
#Aggregating data by 'Country' and Finding sum of 'Sales' 
Country_Sales <- aggregate(Sales ~ Country_Region, data = store_data, sum)

#Changing column name of sales 
colnames(Country_Sales)[2] <- "Total_Sales"

#Finding out Country with highest Sales 
Country_Sales <-arrange(Country_Sales, desc(Total_Sales)) #Arranged Continents based on Sales in descending order

# Converting Country_Region column into factor so that order won't change for graph 
Country_Sales$Country_Region <- factor(Country_Sales$Country_Region, levels = unique(Country_Sales$Country_Region))

#Plotting Country_Region vs TotalSales

#options(repr.plot.width = 30, repr.plot.height = 20)

ggplot(data = Country_Sales, aes(x = Country_Region, y = Total_Sales)) + 
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 10)) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) + ggtitle('Country Regions vs Sales') +
  xlab("Country Regions") + ylab("Total Sales")
```

### Identifying Most Ordered Categories

```{r}
categories <- store_data %>% 
  group_by(Category) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))


ggplot(data = categories, aes(x = reorder(Category, count), y = count))+
  geom_bar(stat = "identity", fill = "#3F6E9A", colour = "#3F6E9A") +
  labs(x = "", y = "Best Sellers", title = "Most Ordered Categories") +
  coord_flip() +
  theme(text = element_text(size = 13))
```

### Identifying most ordered Brands

```{r}
brands <- store_data %>% 
  group_by(Brand) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))


ggplot(data = brands[0:10, ], aes(x = reorder(Brand, count), y = count))+
  geom_bar(stat = "identity", fill = "#3F6E9A", colour = "#3F6E9A") +
  labs(x = "", y = "Top 10 Best Sellers", title = "Most Ordered Brands") +
  coord_flip() +
  theme(text = element_text(size = 13))
```

### Identifying Least ordered Brands

```{r}
ggplot(data = tail(brands, n = 10), aes(x = reorder(Brand, -count), y = count))+
  geom_bar(stat = "identity", fill = "#3F6E9A", colour = "#3F6E9A") +
  labs(x = "", y = "Least Sellers", title = "Least Ordered Brands") +
  scale_y_continuous(limits = c(0, 4), breaks = c(0, 1, 2, 3, 4)) + 
  coord_flip() +
  theme(text = element_text(size = 13))
```

### Identifying most ordered Products

```{r}
# store_data_cleaned <- as.data.frame(gsub("[[:punct:]]", "", as.matrix(store_data)))
products <- store_data %>% 
  group_by(Item) %>% 
  summarize(count = n()) %>%
  arrange(desc(count))


ggplot(data = products[0:10, ], aes(x = reorder(Item, count), y = count))+
  geom_bar(stat = "identity", fill = "#3F6E9A", colour = "#3F6E9A") +
  labs(x = "", y = "Top 10 Best Sellers", title = "Most Ordered Products") +
  coord_flip() +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + 
  theme_grey(base_size = 10) +
  theme(text = element_text(size = 13))
```
```{r}
ggplot(data = tail(products, n = 10), aes(x = reorder(Item, -count), y = count))+
  geom_bar(stat = "identity", fill = "#3F6E9A", colour = "#3F6E9A") +
  labs(x = "", y = "Least Sellers", title = "Least Ordered Products") +
  scale_y_continuous(limits = c(0, 4), breaks = c(0, 1, 2, 3, 4)) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + 
  coord_flip() +
  theme_grey(base_size = 8) +
  theme(text = element_text(size = 13))
```

### Frequency of orders on different week days

```{r}
store_data %>% 
  ggplot(aes(wday(Order_Date, week_start = getOption("lubridate.week.start", 1)))) + 
  geom_histogram(stat = "count" , fill = "forest green", colour = "dark green") +
  labs(x = "Day of Week", y = "") +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7), labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  theme_grey(base_size = 14)
```

### Relationships among numerical variables

```{r}
cordata = store_data[,c(19, 20, 21, 22, 23)]
corr <- round(cor(cordata), 1)
corr
```

#### The output above shows the presence of strong linear correlation between the variables Profit and Sales

```{r}
ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE, lab_size = 3, method="circle", colors = c("blue", "white", "red"), outline.color = "gray", show.legend = TRUE, show.diag = FALSE, title="Correlogram of variables")
```

## Modeling

### RFM Analysis

#### Recency: difference between the analysis date and the most recent date, that the customer has shopped in the store, that is how recently a customer has purchased.

#### Frequency : Number of transactions performed by every customer, that is how often they purchase.

#### Monetary: Total money spent by every customer in the store, that is how much the customer spends.

```{r}
store_data$Order_Date <- as.Date(store_data$Order_Date)
class(store_data$Order_Date)

analysis_date <- max(store_data$Order_Date)
analysis_date
```

```{r}
rfm_df <- store_data %>% group_by(Customer_Id) %>% 
  summarise(Recency = as.numeric(analysis_date - max(Order_Date)), 
            Frequency = n(), Monetary = sum(Total_Price))
rfm_df
```
```{r}
r <- ggplot(rfm_df) + geom_density(aes(x = Recency), color="darkblue", fill="lightblue") + labs(x="Recency \n(How recently customer has purchased)")
f <- ggplot(rfm_df) + geom_density(aes(x = Frequency), color="darkblue", fill="lightblue") + labs(x="Frequency \n(How often customers purchase)")
m <- ggplot(rfm_df) + geom_density(aes(x = Monetary), color="darkblue", fill="lightblue") + labs(x="Monetary \n(How much customers spend)")
grid.arrange(r, f, m, nrow = 3)
```

```{r}
summary(rfm_df)
```

```{r}
rfm_df$R_score <- 0
rfm_df$R_score[rfm_df$Recency >= 539.0] <- 1
rfm_df$R_score[rfm_df$Recency >= 241.0 & rfm_df$Recency < 539.0] <- 2
rfm_df$R_score[rfm_df$Recency >= 99.0 & rfm_df$Recency < 241.0] <- 3
rfm_df$R_score[rfm_df$Recency < 99.0] <- 4

rfm_df$F_score<- 0
rfm_df$F_score[rfm_df$Frequency >= 6.000] <- 4
rfm_df$F_score[rfm_df$Frequency < 6.000 & rfm_df$Frequency >= 4.000] <- 3
rfm_df$F_score[rfm_df$Frequency < 4.000 & rfm_df$Frequency >= 1.000] <- 2
rfm_df$F_score[rfm_df$Frequency < 1.000] <- 1

rfm_df$M_score <- 0
rfm_df$M_score[rfm_df$Monetary >= 7262] <- 4
rfm_df$M_score[rfm_df$Monetary < 7262 & rfm_df$Monetary >= 1987] <- 3
rfm_df$M_score[rfm_df$Monetary >= 417 & rfm_df$Monetary < 1987] <- 2
rfm_df$M_score[rfm_df$Monetary < 417] <- 1

rfm_df <- rfm_df %>% mutate(RFM_score = 100 *R_score +10 * F_score + M_score)

rfm_df
```

```{r}
rfm_df$Segment <- "0"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(444,434,443, 344, 442, 244, 424, 441  ))] <-"Loyalists"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(332,333,342, 343, 334, 412,413,414,431,432,441,421,422,423, 424, 433 ))] <- "Potential Loyalists"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(233,234, 241,311, 312, 313,314,321,322,323,324, 331,  341))] <- "Promising"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(124, 133, 134, 142, 143, 144, 214,224,234, 242, 243, 232 ))] <- "Hesitant"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(122, 123,131 ,132, 141, 212, 213, 221, 222, 223, 231 ))] <- "Need attention"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(111, 112, 113, 114, 121, 131, 211, 311, 411 ))] <-"Detractors"

rfm_df
```

```{r}
table(rfm_df$Segment)
```
```{r}
ggplot(rfm_df) + geom_bar(aes(x = Segment, fill = Segment))+theme(axis.text.x=element_text(angle=90,hjust=1)) +labs(title = "Barplot for Segments of customers")
```
```{r}
x <- table(rfm_df$Segment)
piepercent<- round(100*x/sum(x), 1)
lbls = paste(names(x), piepercent,"%")

plotrix::pie3D(x, main = "Pie chart for Customer Segments", 
                      labels=lbls, labelcex=1, explode=0.05, 
                      theta=1, radius=0.8, start=2, col = hue_pal()(6))

```

#### Loyalists : These are the champion customers who love to buy your brand. They should be suitably rewarded so they remain promoters of the brand.

#### Potential Loyalists : They like your products but probably don’t love it. Persuade them to spend more money with you by coming up with attractive offers and festive discounts.

#### Promising : They have had a good purchase history but not great lately. Ask for feedback, offer the best customer experience. Provide personalized digital marketing to better engage with the customers.

#### Hesitant : These customers are passive customers. Assess feedback from the customers through online/physical surveys. Try to understand their expectations and exceed it every time.

#### Need attention : If no suitable action is taken, these customers have a risk of becoming the detractors. We must prevent this by understanding the needs of this segment and offer them tailor made offers/discounts to make them excited about our brand.

#### Detractors : There is a chance that these customers have been disappointed with the product or brand in the past and no longer want to come back. Detractors have a high probability of giving bad reviews and bad-mouthing the product. Strategize the actions to follow up with the detractors. Address any concerns, make improvements, and offer rectification, if possible.



### Cohort Analysis

#### https://towardsdatascience.com/cohort-analysis-in-r-the-easy-way-424f19a37d18

#### Cohort - a group of people banded together or treated as a group

#### Visualising customer and user retention is a useful way for e.g. retail and subscription businesses to keep track of how long customers and users tend to stay with them and spot differences in how cohort sizes change over time.

#### The cohorts package currently allows users to aggregate by month and day, we will aggregate customer orders by month.

### Creating a wide cohort table

```{r}
store_data %>% cohort_table_month(id_var = Customer_Id, date = Order_Date)
```

#### The above output where rows represent cohorts and columns represent months. In this example, customers belonging to cohort 1 made their first purchase in Jan 2010 and customers belonging to cohort 2 placed their first order in Feb 2010 and so forth.

### Getting percentages

```{r}
store_data %>%
   cohort_table_month(id_var = Customer_Id, date = Order_Date) %>%
   cohort_table_pct()
```

#### As is apparent from the output above, cohorts are by definition complete in the initial month, which of course entails a starting value of 100 %. As time goes by, we should expect to see fewer and fewer of the same customers placing new orders in the months that follow their initial orders.


### Left-shifting the cohort table

#### In order to produce meaningful plots, the values will need to be shifted to the left as an intermediate step.

```{r}
store_data %>%
  cohort_table_month(id_var = Customer_Id,
                       date   = Order_Date) %>%
  cohort_table_pct() %>%
  shift_left()

store_data %>%    
  cohort_table_month(id_var = Customer_Id, date   = Order_Date) %>%
  shift_left_pct()
```

### Converting to a long and plot-friendly format

```{r}
store_data_long <- store_data %>%
   cohort_table_day(Customer_Id, Order_Date) %>%
   shift_left_pct() %>%
   pivot_longer(-cohort) %>%
   mutate(time = as.numeric(str_remove(name,"t")))

store_data_long
```

### Creating line plots

```{r}
store_data_long[store_data_long == 0] <- NA
store_data_long
```


```{r}
store_data_long %>%
  filter(cohort <= 7, time > 0, value > 0) %>%
  ggplot(aes(time, value, colour = factor(cohort), group = cohort)) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  theme_minimal()
```

### Creating a table visualization

```{r}
store_data_long %>%
   filter(time > 0, value > 0) %>%
   ggplot(aes(time, reorder(cohort, desc(cohort)))) +
   geom_raster(aes(fill = log(value))) +
   coord_equal(ratio = 1) +
   geom_text(aes(label = glue::glue("{round(value,0)}%")),
                 size  = 3,
                 colour = "snow") +
  scale_fill_gradient(guide = "none") +
  theme_minimal(base_size = 16) +
  theme(panel.grid   = element_blank(),
        panel.border = element_blank()) +
  labs(y = "cohort")

```
